MLH1 Data Report

### 2019-10-10 doc/Main_Report.rmd
more updates to mixed models, learned my lesson for ordered factors

> *-Which strain has highest HetC?*

@@

## Explaination for the XX and XY differences

@@

## Difference in proportion of chrm classes

@@

## Mixed Model, compare within and between subspecies patterns

@@

```{r MxMdl.DF.setup, echo=FALSE, warning=FALSE}
library(lme4)
library(nlme)
# seperate the data, both sexes, only Mus
# decide on Musc - Molossinus
AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x

@@

#remove Castaneus (not females)
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp.x != "Cast",]
#remove musc strains
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]

#remove PERC, MOLF, CZECH,
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "MOLF",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]

#change molossinus to it's own subsp (grepl)

DF.HetC.MixedModel2 <- DF.HetC.MixedModel

@@

DF.HetC.MixedModel2$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel2$strain), "Mol", 
        ifelse(grepl("MOLF", DF.HetC.MixedModel2$strain), "Mol", DF.HetC.MixedModel2$subsp))

DF.HetC.MixedModel2$subsp <- ifelse(grepl(1, DF.HetC.MixedModel2$subsp), "Dom", 
        ifelse(grepl(2, DF.HetC.MixedModel2$subsp), "Musc", DF.HetC.MixedModel2$subsp))

@@

#pull out values / other things
summary(model.ff2)# indeces for this object: [1]-structure
ranef(model.ff2)
#after coding subsp to chr from factor, the co-efficients are

#intercept should be the Dom female mean, looks right
mean(DF.HetC.MixedModel2$mean_co[(DF.HetC.MixedModel2$sex == "female" &
                                    DF.HetC.MixedModel2$subsp == "Dom" )] ) #26.5
#male effect is -2
#Musc and mol effec is almost -1

@@

DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)

@@

colnames(DF.HetC.MixedModel)
DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
                                                            subsp.y,subsp.x,
                                                            category.y, category.x,
                                                            strain.x,strain.y, mouse.1  )  )
str(DF.HetC.MixedModel)

#remove ncells <10

DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$age.weeks < 25,]

#pooled.cells

DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]

DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
#remove musc strains
DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "AST",]
DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "TOM",]

@@

`r table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain)`

```{r mixed.model,make, echo=FALSE, warning=FALSE}
#build mixed model
#run simulations for accounting for sampling error
# old code uses base plot
library(nlme)
library(lme4)

#L,Q,C  after the factors in the summary mean lme doesn't believe you used ordered factors

#redo the orders just to be sure
#sex, strain, subsp,

DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = TRUE, levels =c("female", "male") )

DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = TRUE, levels =c( "Dom", "Musc") )

DF.HetC.MixedModel$strain<- factor(DF.HetC.MixedModel$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW",
                                                  "PWD", "MSM", "SKIVE", "KAZ") )

#dframe$strain<- factor(dframe$strain, ordered = TRUE, levels =c( "WSB", "G", "LEW", "PERC",
#                                                  "PWD", "MSM", "MOLF","SKIVE", "KAZ", "TOM", "AST","CZECH",
#                                                  "CAST", "HMI",
#                                                  "SPRET", "SPIC", "CAROLI", "F1", "other") )

DF.HetC.MixedModel$mean_co <- as.integer(DF.HetC.MixedModel$mean_co)

#female Dom are the first levels:
#Intercept, 25.767065 mean for Dom female -- NS
#
#make sure the default (ordered factors of strains is set)
#check level order etc.
#build initialize model
model.ff <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )

#pull out values / other things
summary(model.ff)# indeces for this object: [1]-structure
ranef(model.ff)

@@

#pull out values / other things
summary(model.pooled)# indeces for this object: [1]-structure
ranef(model.pooled)
```

pooled cell data collection.

`r table(DF.HetC.MixedModel.pooled$sex, DF.HetC.MixedModel.pooled$strain)`
KAZ, LEWES has ~100 more for males, PWD has ~200 for females.

>>>>

@@

### Explaination for the XX and XY differences

@@

### Difference in proportion of chrm classes

@@

### Mixed Model, compare within and between subspecies patterns

@@

```{r MxMdl.DF.setup, echo=FALSE, warning=FALSE, eval=TRUE}
library(lme4)
library(nlme)
# seperate the data, both sexes, only Mus
# decide on Musc - Molossinus
AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x

@@

#change molossinus to it's own subsp (grepl)
DF.HetC.MixedModel2 <- DF.HetC.MixedModel

@@

@@

#pull out values / other things
#summary(model.ff2)# indeces for this object: [1]-structure
#ranef(model.ff2)
#after coding subsp to chr from factor, the co-efficients are

#intercept should be the Dom female mean, looks right
mean(DF.HetC.MixedModel2$mean_co[(DF.HetC.MixedModel2$sex == "female" &
                                    DF.HetC.MixedModel2$subsp == "Dom" )] ) #26.5

@@

DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)
#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)

@@

#colnames(DF.HetC.MixedModel)
DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
                                                            subsp.y,subsp.x,
                                                            category.y, category.x,
                                                            strain.x,strain.y, mouse.1  )  )

#pooled.cells
DF.HetC.MixedModel.pooled <- MLH1_data[MLH1_data$species == "M.musculus",]
DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$subsp != "Cast",]
#remove musc strains
DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "AST",]
DF.HetC.MixedModel.pooled <- DF.HetC.MixedModel.pooled[DF.HetC.MixedModel.pooled$strain != "TOM",]

@@

`r kable(table(DF.HetC.MixedModel$sex, DF.HetC.MixedModel$strain))`

# Building Mixed Models

```{r mixed.model.make, echo=FALSE, warning=FALSE, eval=FALSE, message=FALSE}
#build mixed model
#run simulations for accounting for sampling error
# old code uses base plot
library(nlme)
library(lme4)

#L,Q,C  appear for ordered factors in the summary mean lme doesn't believe you used ordered factors
#change all the DFs / variables to UNORDERED factors!
#redo the orders just to be sure
#sex, strain, subsp,

DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE, levels =c("female", "male") )

DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE, levels =c( "Dom", "Musc") )

DF.HetC.MixedModel$strain<- factor(DF.HetC.MixedModel$strain, ordered = FALSE, levels =c( "WSB", "G", "LEW",
                                                  "PWD", "MSM", "SKIVE", "KAZ") )

@@

#pull out values / other things
summary(model.pooled)# indeces for this object: [1]-structure
ranef(model.pooled)

model.pooled$coefficients
```

List of mixed models with mean_co

1. mean_co --DF.HetC.MixedModel
2. mean_co.chr  variables coded as factors (for ggplots), how is this changed when they are coded as chr
3. mean_co.fac -- non-ordered factors
4. mean_co.HQ
5. mean_co.HQ.mol
6. Sub-sampling for equal numbers

```{r mixed.models.setup, echo=FALSE}
#make are the DFs for the models

#main file, make everything unordered factors
AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]

DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
#remove musc strains
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "MOLF",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]

DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
                                                            subsp.y,subsp.x,
                                                            category.y, category.x,
                                                            strain.x,strain.y, mouse.1  )  )

DF.HetC.MixedModel$subsp <- factor(DF.HetC.MixedModel$subsp, ordered = FALSE )
DF.HetC.MixedModel$sex <- factor(DF.HetC.MixedModel$sex, ordered = FALSE )
DF.HetC.MixedModel$strain <- factor(DF.HetC.MixedModel$strain, ordered = FALSE )

#main df with variables coded as chr
DF.HetC.MixedModel.chr <- DF.HetC.MixedModel
DF.HetC.MixedModel.chr$subsp <- as.character(DF.HetC.MixedModel$subsp)
DF.HetC.MixedModel.chr$sex <- as.character(DF.HetC.MixedModel$sex)
DF.HetC.MixedModel.chr$strain <- as.character(DF.HetC.MixedModel$strain)

#mean_co.HQ
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]

#mean_co.HQ.mol
DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ

DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl("MSM", DF.HetC.MixedModel.HQ.mol$strain), "Mol", 
        ifelse(grepl("MOLF", DF.HetC.MixedModel.HQ.mol$strain), "Mol", DF.HetC.MixedModel.HQ.mol$subsp))

DF.HetC.MixedModel.HQ.mol$subsp <- ifelse(grepl(1, DF.HetC.MixedModel.HQ.mol$subsp), "Dom", 
        ifelse(grepl(2, DF.HetC.MixedModel.HQ.mol$subsp), "Musc", DF.HetC.MixedModel.HQ.mol$subsp))
DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ.mol[!(is.na(DF.HetC.MixedModel.HQ.mol$mouse) | DF.HetC.MixedModel.HQ.mol$mouse==""), ]

#mean_co.chr.HQ.mol
DF.HetC.MixedModel.chr.HQ.mol <- DF.HetC.MixedModel.HQ.mol

DF.HetC.MixedModel.chr.HQ.mol$subsp <- as.character(DF.HetC.MixedModel.chr.HQ.mol$subsp)
DF.HetC.MixedModel.chr.HQ.mol$sex <- as.character(DF.HetC.MixedModel.chr.HQ.mol$sex)
DF.HetC.MixedModel.chr.HQ.mol$strain <- as.character(DF.HetC.MixedModel.chr.HQ.mol$strain)

```{r MixedModels.stff, echo=FALSE}
#main mean.CO
HetC.MixedModel.mean_co <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel, random=list(strain=pdDiag(~sex) ) )

#mean.co.chr
HetC.MixedModel.mean_co.chr <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.chr, random=list(strain=pdDiag(~sex) ) )

#mean.co.HQ
HetC.MixedModel.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random=list(strain=pdDiag(~sex) ) )

#mean.CO.HQ.mol
HetC.MixedModel.HQ.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ.mol, random=list(strain=pdDiag(~sex) ) )

#mean.HQ.mol.chr
HetC.MixedModel.chr.HQ.mol <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.chr.HQ.mol, random=list(strain=pdDiag(~sex) ) )

The intercept level is Dom female. These are the coefficients

`r kable(HetC.MixedModel.mean_co$coefficients$fixed, col.names = "coefficients")`

`r kable(ranef(HetC.MixedModel.mean_co) )`

changed all to unordered factors.

all the variables are coded as ordered factors.

When the variables are kept factors, and unordered the names appear! yay!

`r kable(HetC.MixedModel.mean_co.chr$coefficients$fixed, caption="Chr coded Model", col.names = "coefficients")`

`r kable(ranef(HetC.MixedModel.mean_co.chr) )`

The variable names now appear. The values are now quite different. Male, sex coefficient is negative. The fixed sex coefficient is the sex effect independant of subspecies or strain.

`r kable(HetC.MixedModel.HQ$coefficients$fixed, caption="mean_co HQ model", col.names = "coefficients")`
`r kable(ranef(HetC.MixedModel.HQ) )`

The factor coded DF the more high quality cells.

`r kable(HetC.MixedModel.HQ.mol$coefficients$fixed, caption="Main model with molossinus", col.names = "coefficient")`
`r kable(ranef(HetC.MixedModel.HQ.mol) )`
With adding molossinus, the subsp names are listed.

`r kable(HetC.MixedModel.chr.HQ.mol$coefficients$fixed, caption="Main model, chr coded with molossinus", col.names = "coefficients")`
`r kable(ranef(HetC.MixedModel.chr.HQ.mol) )`

When the variables/levels are coded as chr, the interaction effects get larger and the other subspecies get smaller. The base sex effect is also larger. The subsp effect is relative to Dom female,

I don't think the subspecies and strains should be order, (maybe factors still makes sense.)

This table doesn't display well

`r table(DF.HetC.MixedModel.pooled$sex,DF.HetC.MixedModel.pooled$strain)`

KAZ, LEWES has ~100 more for males, PWD has ~200 for females.
